---
title: "Week 9 - Homework"
author: "STAT 420, Summer 2018, Ren Bettendorf, renb2"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
urlcolor: cyan
---

***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
```

## Exercise 1 (`longley` Macroeconomic Data)

The built-in dataset `longley` contains macroeconomic data for predicting employment. We will attempt to model the `Employed` variable.

```{r, eval=FALSE}
View(longley)
?longley
```

**(a)** What is the largest correlation between any pair of predictors in the dataset?
```{r}
round(cor(longley), 2)
```
<br />
The largest correlation between any pair of predictors is `Year` and `GNP` with a value of $1$.
<br />
<br />
**(b)** Fit a model with `Employed` as the response and the remaining variables as predictors. Calculate and report the variance inflation factor (VIF) for each of the predictors. Which variable has the largest VIF? Do any of the VIFs suggest multicollinearity?
```{r message=FALSE, warning=FALSE}
emp_model = lm(Employed ~ ., data = longley)
library(faraway)
vif(emp_model)
```
<br />
The largest VIF is `GNP`. Using the typical value of $5$ for a large multicollinearity we see that `GNP.deflator`, `GNP`, `unemployed`, `Population`, and `Year`.
<br />
<br />
**(c)** What proportion of the observed variation in `Population` is explained by a linear relationship with the other predictors?
```{r}
pop_model = lm(Population ~ . - Employed, data = longley)
prop = summary(pop_model)$r.squared
```
<br />
The proportion of observed variation in `Population` is $`r prop`$.
<br />
<br />
**(d)** Calculate the partial correlation coefficient for `Population` and `Employed` **with the effects of the other predictors removed**.
```{r}
emp_cor_model = lm(Employed ~ . - Population, data = longley)
cor_value = cor(resid(pop_model), resid(emp_cor_model))
```

<br />
The partial correlation coefficient for `Population` and `Employed` is  $`r cor_value`$.
<br />
<br />
**(e)** Fit a new model with `Employed` as the response and the predictors from the model in **(b)** that were significant. (Use $\alpha = 0.05$.) Calculate and report the variance inflation factor for each of the predictors. Which variable has the largest VIF? Do any of the VIFs suggest multicollinearity?
```{r}
emp_small_model = lm(Employed ~  Unemployed + Armed.Forces + Year, data = longley)
vif(emp_small_model)
```

<br />
By looking at the summary of the model in **(b)** we see that we would use `Unemployed`, `Armed.Forces`, and `Year` thus leading to the largest VIF is `Year` and none of these suggestion multicollinearity.
<br />
<br />
**(f)** Use an $F$-test to compare the models in parts **(b)** and **(e)**. Report the following:

- The null hypothesis
- The test statistic
- The distribution of the test statistic under the null hypothesis
- The p-value
- A decision
- Which model you prefer, **(b)** or **(e)**
```{r}
anova_emp = anova(emp_model, emp_small_model)
```
<br />
 
  - Null Hypothesis: $H_0: \beta_{GNP.Deflator} = \beta_{GNP} = \beta_{Population} = 0$
  - Alternative Hypothesis: $H_1: \beta_{GNP.Deflator}, \beta_{GNP}, \beta_{Population} \neq 0$
  - Test Statistic: $`anova_emp[2, 5]`$.
  - Distribution of the test satistic is F with DF 12 and 9.
  - P-Value: $`r anova_emp[2, 6]`$.
  - Decision: Fail to Reject $H_0$.
  - Model Preference: We prefer the smaller model using `Unemployed`, `Armed.Forces`, and `Year`.
  
<br />
<br />
**(g)** Check the assumptions of the model chosen in part **(f)**. Do any assumptions appear to be violated?

```{r, echo = FALSE}
plot_fitted_resid = function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  plot(fitted(model), resid(model), 
       col = pointcol, pch = 20, cex = 1.5,
       xlab = "Fitted", ylab = "Residuals")
  abline(h = 0, col = linecol, lwd = 2)
}

plot_qq = function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  qqnorm(resid(model), col = pointcol, pch = 20, cex = 1.5)
  qqline(resid(model), col = linecol, lwd = 2)
}
```
```{r}
plot_fitted_resid(emp_small_model)
```
```{r}
plot_qq(emp_small_model)
```
```{r}
library(lmtest)
bptest(emp_small_model)
shapiro.test(resid(emp_small_model))$p.value
```
```{r}

```

<br />
Excluding one point we can see that the data is fairly evenly spread out and looking at the QQ plot, we can see that it follows the line mostly, except for a couple of points at the edges so it appears no assumptions have been violated. Furthermore, if we look at the Breusch-Pagan Test we can fail to reject that we do not have constant variance along with the Shapiro-Wilk Test for normality.

***

## Exercise 2 (`Credit` Data)

For this exercise, use the `Credit` data from the `ISLR` package. Use the following code to remove the `ID` variable which is not useful for modeling.

```{r}
library(ISLR)
data(Credit)
Credit = subset(Credit, select = -c(ID))
```

Use `?Credit` to learn about this dataset.

**(a)** Find a "good" model for `balance` using the available predictors. Use any methods seen in class except transformations of the response. The model should:

- Reach a LOOCV-RMSE below `135`
- Obtain an adjusted $R^2$ above `0.90`
- Fail to reject the Breusch-Pagan test with an $\alpha$ of $0.01$
- Use fewer than 10 $\beta$ parameters

Store your model in a variable called `mod_a`. Run the two given chunks to verify your model meets the requested criteria. If you cannot find a model that meets all criteria, partial credit will be given for meeting at least some of the criteria.
```{r}
mod_a = lm(Balance ~ log(Income) + Limit + Education + Married + Education:Student + Married:Age, data = Credit)
```

```{r, message = FALSE, warning = FALSE}
library(lmtest)

get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}
```

```{r}
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```

**(b)** Find another "good" model for `balance` using the available predictors. Use any methods seen in class except transformations of the response. The model should:

- Reach a LOOCV-RMSE below `125`
- Obtain an adjusted $R^2$ above `0.91`
- Fail to reject the Shapiro-Wilk test with an $\alpha$ of $0.01$
- Use fewer than 25 $\beta$ parameters

Store your model in a variable called `mod_b`. Run the two given chunks to verify your model meets the requested criteria. If you cannot find a model that meets all criteria, partial credit will be given for meeting at least some of the criteria.
```{r}
mod_b = lm(Balance ~ Income * Rating * Limit * Student, data = Credit)
```

```{r, message = FALSE, warning = FALSE}
library(lmtest)

get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}
```

```{r}
get_loocv_rmse(mod_b)
get_adj_r2(mod_b)
get_sw_decision(mod_b, alpha = 0.01)
get_num_params(mod_b)
```

***

## Exercise 3 (`Sacramento` Housing Data)

For this exercise, use the `Sacramento` data from the `caret` package. Use the following code to perform some preprocessing of the data.

```{r message=FALSE, warning=FALSE}
library(caret)
library(ggplot2)
data(Sacramento)
sac_data = Sacramento
sac_data$limits = factor(ifelse(sac_data$city == "SACRAMENTO", "in", "out"))
sac_data = subset(sac_data, select = -c(city, zip))
```

Instead of using the `city` or `zip` variables that exist in the dataset, we will simply create a variable (`limits`) indicating whether or not a house is technically within the city limits of Sacramento. (We do this because they would both be factor variables with a **large** number of levels. This is a choice that is made due to laziness, not necessarily because it is justified. Think about what issues these variables might cause.)

Use `?Sacramento` to learn more about this dataset.

A plot of longitude versus latitude gives us a sense of where the city limits are.

```{r}
qplot(y = longitude, x = latitude, data = sac_data,
      col = limits, main = "Sacramento City Limits ")
```

After these modifications, we test-train split the data.

```{r}
set.seed(420)
sac_trn_idx  = sample(nrow(sac_data), size = trunc(0.80 * nrow(sac_data)))
sac_trn_data = sac_data[sac_trn_idx, ]
sac_tst_data = sac_data[-sac_trn_idx, ]
```

The training data should be used for all model fitting. Our goal is to find a model that is useful for predicting home prices.

**(a)** Find a "good" model for `price`. Use any methods seen in class. The model should reach a LOOCV-RMSE below 77,500 in the training data. Do not use any transformations of the response variable.
```{r}
price_model = lm(price ~ beds * sqft * latitude * longitude, data = sac_trn_data)
get_loocv_rmse(price_model)
```

**(b)** Is a model that achieves a LOOCV-RMSE below 77,500 useful in this case? That is, is an average error of 77,500 low enough when predicting home prices? To further investigate, use the held-out test data and your model from part **(a)** to do two things:

- Calculate the average percent error:
\[
\frac{1}{n}\sum_i\frac{|\text{predicted}_i - \text{actual}_i|}{\text{predicted}_i} \times 100
\]
- Plot the predicted versus the actual values and add the line $y = x$.

Based on all of this information, argue whether or not this model is useful.
```{r}
n = nrow(sac_tst_data)
test_actual = sac_tst_data$price
test_predicted = predict(price_model, newdata = data.frame(beds = c(sac_tst_data$beds), sqft = c(sac_tst_data$sqft), latitude = c(sac_tst_data$latitude), longitude = c(sac_tst_data$longitude)))
average_percent_error = (1/n) * sum(abs(test_predicted-test_actual)/test_predicted) *100
```

<br />
The reported Average Percent Error: $`r average_percent_error`$.
<br />
```{r}
plot(test_predicted, test_actual, col= "blue")
abline(0, 1, lwd=2, col="darkorange")
```
<br />
Even though the LOOCV-RMSE is quite high at 76,231 we can see from the graph that it does a fairly good job of predicting the values and dealing with the outliers that would skew the dataset.
<br />
<br />
***

## Exercise 4 (Does It Work?)

In this exercise, we will investigate how well backwards AIC and BIC actually perform. For either to be "working" correctly, they should result in a low number of both **false positives** and **false negatives**. In model selection,

- **False Positive**, FP: Incorrectly including a variable in the model. Including a *non-significant* variable
- **False Negative**, FN: Incorrectly excluding a variable in the model. Excluding a *significant* variable

Consider the **true** model

\[
Y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + \beta_4 x_4 + \beta_5 x_5 + \beta_6 x_6 + \beta_7 x_7 + \beta_8 x_8 + \beta_9 x_9 + \beta_{10} x_{10} + \epsilon
\]

where $\epsilon \sim N(0, \sigma^2 = 4)$. The true values of the $\beta$ parameters are given in the `R` code below.

```{r}
beta_0  = 1
beta_1  = -1
beta_2  = 2
beta_3  = -2
beta_4  = 1
beta_5  = 1
beta_6  = 0
beta_7  = 0
beta_8  = 0
beta_9  = 0
beta_10 = 0
sigma = 2
```

Then, as we have specified them, some variables are significant, and some are not. We store their names in `R` variables for use later.

```{r}
not_sig  = c("x_6", "x_7", "x_8", "x_9", "x_10")
signif = c("x_1", "x_2", "x_3", "x_4", "x_5")
```

We now simulate values for these `x` variables, which we will use throughout part **(a)**.

```{r}
set.seed(19890927)
n = 100
num_sims = 300
x_1  = runif(n, 0, 10)
x_2  = runif(n, 0, 10)
x_3  = runif(n, 0, 10)
x_4  = runif(n, 0, 10)
x_5  = runif(n, 0, 10)
x_6  = runif(n, 0, 10)
x_7  = runif(n, 0, 10)
x_8  = runif(n, 0, 10)
x_9  = runif(n, 0, 10)
x_10 = runif(n, 0, 10)
```

We then combine these into a data frame and simulate `y` according to the true model.

```{r}
sim_data_1 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma)
)
```

We do a quick check to make sure everything looks correct.

```{r}
head(sim_data_1)
```

Now, we fit an incorrect model.

```{r}
fit = lm(y ~ x_1 + x_2 + x_6 + x_7, data = sim_data_1)
coef(fit)
```

Notice, we have coefficients for `x_1`, `x_2`, `x_6`, and `x_7`. This means that `x_6` and `x_7` are false positives, while `x_3`, `x_4`, and `x_5` are false negatives.

To detect the false negatives, use:

```{r}
# which are false negatives?
!(signif %in% names(coef(fit)))
```

To detect the false positives, use:

```{r}
# which are false positives?
names(coef(fit)) %in% not_sig
```

Note that in both cases, you could `sum()` the result to obtain the number of false negatives or positives.

**(a)** Set a seed equal to your birthday; then, using the given data for each `x` variable above in `sim_data_1`, simulate the response variable `y` 300 times. Each time,

- Fit an additive model using each of the `x` variables.
- Perform variable selection using backwards AIC.
- Perform variable selection using backwards BIC.
- Calculate and store the number of false negatives for the models chosen by AIC and BIC.
- Calculate and store the number of false positives for the models chosen by AIC and BIC.

Calculate the rate of false positives and negatives for both AIC and BIC. Compare the rates between the two methods. Arrange your results in a well formatted table.
```{r}
set.seed(19890927)
num_sims = 300
false_negative_aic1 = rep(0, num_sims)
false_negative_bic1 = rep(0, num_sims)
false_positive_aic1 = rep(0, num_sims)
false_positive_bic1 = rep(0, num_sims)

for(i in 1:num_sims) {
  sim_data_1 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
      y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma))
  start_model = lm(y ~ ., data = sim_data_1)
  aic_model = step(start_model, trace = 0)
  n = length(resid(start_model))
  bic_model = step(start_model, k = log(n), trace = 0)
  
  false_negative_aic1[i] = sum(!(signif %in% names(coef(aic_model))))
  false_positive_aic1[i] = sum(names(coef(aic_model)) %in% not_sig)
  false_negative_bic1[i] = sum(!(signif %in% names(coef(bic_model))))
  false_positive_bic1[i] = sum(names(coef(bic_model)) %in% not_sig)
}
```
```{r}
library(knitr)
sim_data_table = data.frame(StepMethod = c("AIC", "BIC"), 
                           Negative = c(sum(false_negative_aic1)/(num_sims * 5), 
                                        sum(false_negative_bic1)/(num_sims * 5)), 
                           Positive = c(sum(false_positive_aic1)/(num_sims * 5), 
                                        sum(false_positive_bic1)/(num_sims * 5)))
kable(sim_data_table)
```

**(b)** Set a seed equal to your birthday; then, using the given data for each `x` variable below in `sim_data_2`, simulate the response variable `y` 300 times. Each time,

- Fit an additive model using each of the `x` variables.
- Perform variable selection using backwards AIC.
- Perform variable selection using backwards BIC.
- Calculate and store the number of false negatives for the models chosen by AIC and BIC.
- Calculate and store the number of false positives for the models chosen by AIC and BIC.

Calculate the rate of false positives and negatives for both AIC and BIC. Compare the rates between the two methods. Arrange your results in a well formatted table. Also compare to your answers in part **(a)** and suggest a reason for any differences.

```{r}
set.seed(19890927)
x_1  = runif(n, 0, 10)
x_2  = runif(n, 0, 10)
x_3  = runif(n, 0, 10)
x_4  = runif(n, 0, 10)
x_5  = runif(n, 0, 10)
x_6  = runif(n, 0, 10)
x_7  = runif(n, 0, 10)
x_8  = x_1 + rnorm(n, 0, 0.1)
x_9  = x_1 + rnorm(n, 0, 0.1)
x_10 = x_2 + rnorm(n, 0, 0.1)

sim_data_2 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma)
)
```
```{r}
set.seed(19890927)
num_sims = 300
false_negative_aic2 = rep(0, num_sims)
false_negative_bic2 = rep(0, num_sims)
false_positive_aic2 = rep(0, num_sims)
false_positive_bic2 = rep(0, num_sims)

for(i in 1:num_sims) {
  sim_data_2 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma))
  start_model = lm(y ~ ., data = sim_data_2)
  aic_model = step(start_model, trace = 0)
  n = length(resid(start_model))
  bic_model = step(start_model, k = log(n), trace = 0)
  
  false_negative_aic2[i] = sum(!(signif %in% names(coef(aic_model))))
  false_positive_aic2[i] = sum(names(coef(aic_model)) %in% not_sig)
  false_negative_bic2[i] = sum(!(signif %in% names(coef(bic_model))))
  false_positive_bic2[i] = sum(names(coef(bic_model)) %in% not_sig)
}
```
```{r}
library(knitr)
sim_data_table = data.frame(StepMethod = c("AIC", "BIC"), 
                           Negative = c(sum(false_negative_aic2)/(num_sims * 5), 
                                        sum(false_negative_bic2)/(num_sims * 5)), 
                           Positive = c(sum(false_positive_aic2)/(num_sims * 5), 
                                        sum(false_positive_bic2)/(num_sims * 5)))
kable(sim_data_table)
```

<br />
As we can see from the first simulation we are not falsely excluding any significant predictors for either AIC or BIC, but it does choose more false positives for AIC over BIC due to AIC prefering a larger model. In the second simulation we see some false negatives due to the collinearity we introduced so the steps are choosing other `x_8`, `x_9`, or `x_10` sometimes due to this and we see more false positives as a result also as it is harder to remove these variables.
<br />