---
title: "Week 3 - Homework"
author: "STAT 420, Summer 2019, Ren Bettendorf"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
urlcolor: cyan
---


## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.
```{r, echo=FALSE}
library(MASS)
```

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
cat_model = lm(Hwt ~ Bwt, data = cats)
```
+ The null and alternative test that we will perform can be written as $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ which means we will check whether $x$, cat body weight, has a linear relationship with $y$, cat heart weight.
   we are specifically testing for $\beta_1 = 0$ and can think of model this way, 
   <br/>
    - Under $H_0$, there is not a significant linear relationship between $x$ and $y$.
   <br/>
    - Under $H_1$, there is a significant **linear** relationship between $x$ and $y$.
    <br />
+ The value of test statistics for $\beta_1$ is $`r summary(cat_model)$coefficients[2,3]`$. 
+ The p-value of the test for $\beta_1$ is $`r summary(cat_model)$coefficients[2,4]`$.
+ Using $\alpha = 0.01$ value compared to p-value of $`r summary(cat_model)$coefficients[2,4]`$, we would reject the null hypothesis thus stating there is a significant linear relationship between cat heart weight and body weight
+ Under the null hypothesis, we assumed no linear relation between cats heart and body weight but using t test and p-value we proved a significant linear relationship between the cats heart and corresponding body weight.
<br />
<br />
**(b)** Calculate a 90% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.
```{r}
confint_b = confint(cat_model, level = 0.90)[2,]
confint_b
```
- This creates a interval where we are 90% sure that the average cat body weight slope resides between $`r confint_b[1]`$ kg and $`r confint_b[2]`$ kg. From part A, we showed we are fairly sure this value is close to the actual.

**(c)** Calculate a 99% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.
```{r}
confint_c = confint(cat_model, level = 0.99)[1,]
confint_c
```
- This creates a interval where we are 90% sure that the average cat heart weight when body weight is $0$ kg resides between $`r  confint_c[1]`$ kg and $`r confint_c[2]`$ kg. The lower estimate of this value is a bit odd as it shows a negative value which shouldn't exist in the real world.

**(d)** Use a 99% confidence interval to estimate the mean heart weight for body weights of $2.1$ and $2.8$ kilograms. Which of the two intervals is wider? Why?
```{r}
new_weights = data.frame(Bwt = c(2.1, 2.8))
predict(cat_model, newdata = new_weights, interval = c("confidence"), level = 0.99)
```
- We can see that the interval of $2.1$ is larger due to being slightly further away from $`r mean(cats$Bwt)`$ which means that our prediction will take this into account when creating the new confidence intervals.

**(e)** Use a 99% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.
```{r}
new_weights = data.frame(Bwt = c(2.8, 4.2))
predict_99 = predict(cat_model, newdata = new_weights, interval = c("prediction"), level = 0.99)
```
- The reported intervals for $2.8$ are $(`r predict_99[1, 1]`, `r predict_99[1, 2]`)$ and for $4.2$ are $(`r predict_99[2, 1]`, `r predict_99[2, 2]`)$

**(f)** Create a scatterplot of the data. Add the regression line, 90% confidence bands, and 90% prediction bands.
```{r}
cat_grid = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)
cat_ci_band = predict(cat_model, newdata = data.frame(Bwt = cat_grid), interval = "confidence", level = 0.90)
cat_pi_band = predict(cat_model, newdata = data.frame(Bwt = cat_grid), interval = "prediction", level = 0.90)
plot(Hwt ~ Bwt, data = cats,
     xlab = "Body Weight (Kg)",
     ylab = "Heart Weight (g)",
     main = "Body weight vs Heart Weight",
     pch  = 20,
     cex  = 1,
     col  = "grey",
     ylim = c(min(cat_pi_band), max(cat_pi_band)))
abline(cat_model, lwd = 3, col = "darkorange")
lines(cat_grid, cat_ci_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(cat_grid, cat_ci_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(cat_grid, cat_pi_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 3)
lines(cat_grid, cat_pi_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 3)
points(mean(cats$Bwt), mean(cats$Hwt), pch = "+", cex = 3)
```
<br />
**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
cat_model_test_info = summary(cat_model)$coefficients
beta_1_hat = cat_model_test_info[2, 1]
beta_1_hat_se = cat_model_test_info[2, 2]
beta_1_hat_t = (beta_1_hat - 4) / beta_1_hat_se
beta_1_hat_pval = 2 * pt(abs(beta_1_hat_t), df = length(resid(cat_model)) - 2, lower.tail = FALSE)
```
+ The value of test statistics for $\beta_1$ is $`r beta_1_hat_t`$. 
+ The p-value of the test for $\beta_1$ is $`r beta_1_hat_pval`$.
+ Using $\alpha = 0.05$ value compared to p-value of $`r beta_1_hat_pval`$, we would fail to reject the null hypothesis thus stating there is a significant chance that a $\beta_1 = 4$ can occur.
<br />
***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
ozone_wind_model = lm(ozone ~ wind, data = Ozone)
```
+ The null and alternative test that we will perform can be written as $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ which means we will check whether $x$, wind speed, has a linear relationship with $y$, ozone.
   we are specifically testing for $\beta_1 = 0$ and can think of model this way, 
   <br/>
    - Under $H_0$, there is not a significant linear relationship between $x$ and $y$.
   <br/>
    - Under $H_1$, there is a significant **linear** relationship between $x$ and $y$.
    <br />
+ The value of test statistics for $\beta_1$ is $`r summary(ozone_wind_model)$coefficients[2,3]`$. 
+ The p-value of the test for $\beta_1$ is $`r summary(ozone_wind_model)$coefficients[2,4]`$.
+ Using $\alpha = 0.01$ value compared to p-value of $`r summary(ozone_wind_model)$coefficients[2,4]`$, we would fail to reject the null hypothesis thus stating there is not a significant linear relationship between wind speed and ozone coverage
+ Under the null hypothesis, we assumed no linear relationship between wind speed and ozone coverage but using t test and p-value we proved that we were unable to reject this hypothesis and show that there most likely there isn't a relationship between the two.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
ozone_temp_model = lm(ozone ~ temp, data = Ozone)
```
+ The null and alternative test that we will perform can be written as $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ which means we will check whether $x$, temperature, has a linear relationship with $y$, ozone.
   we are specifically testing for $\beta_1 = 0$ and can think of model this way, 
   <br/>
    - Under $H_0$, there is not a significant linear relationship between $x$ and $y$.
   <br/>
    - Under $H_1$, there is a significant **linear** relationship between $x$ and $y$.
    <br />
+ The value of test statistics for $\beta_1$ is $`r summary(ozone_temp_model)$coefficients[2,3]`$. 
+ The p-value of the test for $\beta_1$ is $`r summary(ozone_temp_model)$coefficients[2,4]`$.
+ Using $\alpha = 0.01$ value compared to p-value of $`r summary(ozone_temp_model)$coefficients[2,4]`$, we would reject the null hypothesis thus stating there is a significant linear relationship between temperature and ozone coverage.
+ Under the null hypothesis, we assumed no linear relationship between temperature and ozone coverage but using t test and p-value we proved that we were able to reject this hypothesis and show that there most likely there is a linear relationship between the two.
***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19890927
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```
```{r}
beta_0 = -5
beta_1 = 3.25
sigma = 4
Sxx = sum((x - mean(x))^2)
var_beta_1_hat = sigma^2 / Sxx 
var_beta_0_hat = sigma^2 * (1/n + mean(x)^2/Sxx)
beta_0_hats = rep(0, 2000)
beta_1_hats = rep(0, 2000)
for(i in 1:2000) {
  eps = rnorm(n, mean=0, sd=sigma)
  y = beta_0 + beta_1*x + eps
  model = lm(y ~ x)
  beta_0_hats[i] = coef(model)[1]
  beta_1_hats[i] = coef(model)[2]
}
```


**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values
```{r, echo=FALSE}
library(knitr)
df = data.frame(Names=c("True Expected X", "Mean Simulated", "True SD X", "SD Simulated"), 
                beta_0_hat=c(beta_0, mean(beta_0_hats), var_beta_0_hat, sd(beta_0_hats)),
                beta_1_hat=c(beta_1, mean(beta_1_hats), var_beta_1_hat, sd(beta_1_hats)))
kable(df)
```

**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.
```{r}
hist(beta_0_hats, prob = TRUE, breaks = 25, 
     xlab = expression(hat(beta)[0]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = beta_0, sd = sqrt(var_beta_0_hat)),
      col = "darkorange", add = TRUE, lwd = 3)
```

```{r}
hist(beta_1_hats, prob = TRUE, breaks = 20, 
     xlab = expression(hat(beta)[1]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = beta_1, sd = sqrt(var_beta_1_hat)), 
      col = "darkorange", add = TRUE, lwd = 3)
```


***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19890927
set.seed(birthday)
n = 25
x = seq(0, 2.5, length = n)
```
```{r}
beta_0 = 5
beta_1 = 2
sigma = 3

Sxx = sum((x - mean(x))^2)

beta_0_hats = rep(0, 2500)
beta_1_hats = rep(0, 2500)
beta_0_hats_se = rep(0, 2500)
beta_1_hats_se = rep(0, 2500)
for(i in 1:2500) {
  eps = rnorm(n,mean=0,sd=sigma)
  y = beta_0 + beta_1 * x + eps

  model = lm(y ~ x)
  
  beta_0_hats[i] = coef(model)[1]
  beta_1_hats[i] = coef(model)[2]
  beta_0_hats_se[i] = summary(model)$coefficient[1,2]
  beta_1_hats_se[i] = summary(model)$coefficient[2,2]
}
```


**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.
```{r}
critical_value = qt(0.975, df = length(x) - 2)
lower_95 = beta_1 - critical_value * beta_1_hats_se
upper_95 = beta_1 + critical_value * beta_1_hats_se

intervals = data.frame(lower_95 = lower_95, upper_95 = upper_95)
```

**(c)** What proportion of these intervals contains the true value of $\beta_1$?
```{r}
subset_intervals = subset(intervals, lower_95 < beta_1 & beta_1 < upper_95)
proportion = nrow(subset_intervals) / nrow(intervals)
```
Proportion that contains $\beta_1$ is $`r proportion`$.
<br />
**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?
```{r}
subset_intervals = subset(intervals, lower_95 < 0 & 0 < upper_95)
proportion = nrow(subset_intervals) / nrow(intervals)
```
Proportion that contains $\beta_1 = 0$ is $`r proportion`$ thus we fail to reject the null hypothesis as $\alpha = 0.05$.
<br />
**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.
```{r}
critical_value = qt(0.995, df = length(x) - 2)
lower_99 = beta_1 - critical_value * beta_1_hats_se
upper_99 = beta_1 + critical_value * beta_1_hats_se

intervals = data.frame(lower_99 = lower_99, upper_99 = upper_99)
```

**(f)** What proportion of these intervals contains the true value of $\beta_1$?
```{r}
subset_intervals = subset(intervals, lower_99 < beta_1 & beta_1 < upper_99)
proportion = nrow(subset_intervals ) / nrow(intervals)
```
Proportion that contains $\beta_1$ is $`r proportion`$.
<br />
**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?
```{r}
subset_intervals = subset(intervals, lower_99 < 0 & 0 < upper_99)
proportion = nrow(subset_intervals) / nrow(intervals)
```
Proportion that contains $\beta_1 = 0$ is $`r proportion`$ thus we fail to reject the null hypothesis as $\alpha = 0.01$.
<br />
***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval
```{r}
calc_pred_int = function(model, newdata, level = 0.95) {
  estimate = predict(model, newdata = newdata)
  n = length(model$residuals)
  degree_freedom = n - 2
  x = model$model[,2]
  se = summary(model)$sigma
  alpha = 1 - level
  conf = level + alpha/2
  Sxx = sum((x - mean(x))^2)
  rad = 1 + 1/n + (newdata[1] - mean(x))^2/Sxx
  lower = estimate - qt(conf, df = degree_freedom) * se * sqrt(rad)
  upper = estimate + qt(conf, df = degree_freedom) * se * sqrt(rad)
  return(list(estimate=estimate, lower=lower, upper=upper))
}
```


**(b)** After writing the function, run this code:

```{r}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```
**(c)** After writing the function, run this code:

```{r}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.99)
```

